---
title: "260507 MAPT Peptide Level"
output: html_document
date: "2026-05-07"
---
```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
```

Colours for plotting
```{r}
plot_colours <- c("cyan4", "indianred", "brown")
```

Optional ggsave function
```{r}
save_option <- TRUE

ggsave_optional <- function(filename) {
  peptide_analysis_plot_folder <- "./tPx MAPT Peptide Plots/"
  if(save_option){
    ggsave(paste0(peptide_analysis_plot_folder, filename))
  } else {
    print("Plot not saved")
  }
}
```

Load pre-filtered MAPT peptide dataset
```{r}
MAPTpeptides <- read.csv("All_MAPT_Peptides.csv")
MAPTpeptides <- MAPTpeptides[,2:4]
```

Bring in metadata and tau441 peptide sequence
```{r}
mapper <- read.csv("totalproteome_mapper.csv")
Tau441 <- toupper(as.character(read.delim("Tau441 Amino Acid Sequence.txt", header = FALSE)))

```

Prepare data
```{r} 
#reshape to wide format
MAPTpeptidesWide <- suppressWarnings(reshape(MAPTpeptides, idvar = "Stripped.Sequence", timevar = "Run", direction = "wide"))
#replace column names with Simple IDs
colnames(MAPTpeptidesWide)[2:21] <- sapply(strsplit(colnames(MAPTpeptidesWide[2:21]),"_"), `[`, 6)

#make column with peptide start position
MAPTpeptidesWide$PeptideStart <- sapply(
  MAPTpeptidesWide$Stripped.Sequence,
  function(x) regexpr(x, Tau441, fixed = TRUE)[1]
)

#make column with peptide end position (position of the final amino acid of the peptide), and then sort the data by peptide position 
MAPTpeptidesWide <- MAPTpeptidesWide %>% mutate(PeptideEnd = PeptideStart + nchar(Stripped.Sequence) - 1) %>% relocate(PeptideStart, .after = Stripped.Sequence) %>% relocate(PeptideEnd, .after = PeptideStart) %>% arrange(PeptideStart, PeptideEnd)
```

```{r}
#Count NAs per peptide
NACountByRow <- rowSums(is.na(MAPTpeptidesWide))
table(NACountByRow)

#Filter by number of allowed NAs (out of 20 samples)
MaxNAs <- 6
MAPTpeptidesWide <- MAPTpeptidesWide[NACountByRow<=MaxNAs,]
```

```{r}
#filter out unidentified peptides (exon junctions not in tau441 probably? - Actually can't seem to find at all. not sure what these are? Possible in some non-neuronal tau exon than isn't present in tau441 at all, e.g. ex4a)
MAPTpeptidesWide <- MAPTpeptidesWide %>% filter(PeptideStart>0)

#remove ex9-ex10 peptide that includes N279K mutation site
MAPTpeptidesWide <- MAPTpeptidesWide %>% filter(Stripped.Sequence!="HQPGGGKVQIINK")

#Add in a PeptideRank column to help with plotting later
MAPTpeptidesWide$PeptideRank <- rank(MAPTpeptidesWide$PeptideStart, ties.method = "first")
MAPTpeptidesWide <- MAPTpeptidesWide %>% relocate(PeptideRank, .after = Stripped.Sequence)
```

Make a new column for Exon 10 status
```{r}
MAPTpeptidesWide$InExon10 <- 
  (MAPTpeptidesWide$PeptideStart<=305 & MAPTpeptidesWide$PeptideStart>=275)|
  (MAPTpeptidesWide$PeptideEnd<=305 & MAPTpeptidesWide$PeptideEnd>=275)
MAPTpeptidesWide <- MAPTpeptidesWide %>% relocate(InExon10, .after = PeptideEnd)
```


Normalisation
```{r}
#normalise within each sample - NOW ONLY USING PEPTIDES FOR CONSTITUTIVE EXONS
SampleMeans <- colMeans(MAPTpeptidesWide[!MAPTpeptidesWide$InExon10,6:25], na.rm = TRUE)
SampleSds <- apply(MAPTpeptidesWide[!MAPTpeptidesWide$InExon10,6:25], 2, sd, na.rm = TRUE)

MAPTpeptidesWide[,6:25] <- scale(MAPTpeptidesWide[,6:25], center = SampleMeans, scale = SampleSds)

#initial col means and sds (excluding E10 peptides)
print("======== Old Sample Means ========")
print(SampleMeans[1:5])
print("======== Old Sample SDs ========")
print(SampleSds[1:5])

#new col means and sds (excluding E10 peptides)
print("======== New Sample Means ========")
print(round(colMeans(MAPTpeptidesWide[!MAPTpeptidesWide$InExon10,6:25], na.rm = TRUE), digits = 3)[1:5])
print("======== New Sample SDs ========")
print(round(apply(MAPTpeptidesWide[!MAPTpeptidesWide$InExon10,6:25], 2, sd, na.rm = TRUE), digits = 3)[1:5])

```
Peptide-wise normalisation
```{r}
PeptideWiseScaled <- MAPTpeptidesWide
PeptideWiseScaled[,6:25] <- t(scale(t(PeptideWiseScaled[,6:25])))
```

Melt to long format for plotting
```{r}
MAPTpeptidesMelt <- melt(MAPTpeptidesWide, idvars = "Stripped.Sequence", measure.vars = as.character(c(1:20)))
names(MAPTpeptidesMelt)[names(MAPTpeptidesMelt) == "variable"] <- "ID"
head(MAPTpeptidesMelt)

#Add Genotype column
mapper$ID <- as.factor(mapper$ID)
MAPTpeptidesMelt <- left_join(MAPTpeptidesMelt, mapper[,colnames(mapper) %in% c("ID", "Genotype")], by = "ID")

```

Summary stats by genotype
```{r}
GenotypeStats <- MAPTpeptidesMelt %>%
  group_by(Genotype, Stripped.Sequence) %>%
  summarise(
    count = n(),
    mean_val = mean(value, na.rm = TRUE),
    sd_val = sd(value, na.rm = TRUE)
  )

GenotypeStats <- left_join(GenotypeStats, MAPTpeptidesWide[,1:5], by = "Stripped.Sequence")
head(GenotypeStats)
```
Exon 10 data and peptide ranks/positions
```{r}
GenotypeStatsE10 <- GenotypeStats[GenotypeStats$InExon10,]
E10MinRank <- min(GenotypeStatsE10$PeptideRank)
E10MaxRank <- max(GenotypeStatsE10$PeptideRank)
E10MinPos <- min(GenotypeStatsE10$PeptideStart)
E10MaxPos <- max(GenotypeStatsE10$PeptideStart)

```


Plot by genotype
```{r, fig.width=10,fig.height=6.5}
p <- ggplot(data=GenotypeStats, aes(x=PeptideRank, y=mean_val, group = Genotype)) +
  geom_rect(ymax = Inf, ymin = -Inf, xmin = E10MinRank-0.5, xmax = E10MaxRank+0.5, fill = alpha("darkgoldenrod1", 0.005)) +
  geom_text(
    label="Exon 10", 
    x=mean(c(E10MinRank,E10MaxRank)),
    y=4.8,
    color = "#002000",
    aes(fontface=2),
    size = 5
  ) +
  geom_line(aes(color = Genotype), linewidth = 1.5, lineend = "round")+
  geom_point(aes(color = Genotype)) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val, color = Genotype), linewidth = 1, width = 0.4) +
  ggtitle("Peptides Ranked by Start Position") + 
  scale_color_manual(values=plot_colours) + 
  theme_classic() +
  ylab("Peptide Abundance\n(standardised against the mean of constitutive peptides)") +
  xlab("Peptide (ranked by start position)")
  #facet_wrap to separate genotypes, remove to overlay as default
  #facet_wrap(~Genotype, ncol=3)
print(p)

ggsave_optional("Unscaled all peptides lineplot by rank.png")

q <- ggplot(data=GenotypeStats, aes(x=PeptideStart, y=mean_val, group = Genotype)) +
  geom_rect(ymax = Inf, ymin = -Inf, xmin = E10MinPos-5, xmax = E10MaxPos+5, fill = alpha("darkgoldenrod1", 0.005)) +
  geom_text(
    label="Exon 10", 
    x=mean(c(E10MinPos, E10MaxPos))-5,
    y=4.8,
    color = "#002000",
    aes(fontface=2),
    size = 5
  ) +
  geom_line(aes(color = Genotype), linewidth = 1.5, lineend = "round")+
  geom_point(aes(color = Genotype)) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val, color = Genotype), linewidth = 1, width = 3) +
  ggtitle("Peptides by Actual Start Position") + 
  scale_color_manual(values=plot_colours) +
  theme_classic() +
  ylab("Peptide Abundance\n(standardised against the mean of constitutive peptides)") +
  xlab("Peptide Start Position") 
print(q)

ggsave_optional("Unscaled all peptides lineplot by position.png")
```

Bar plot E10 only
```{r}
GenotypeStatsE10 <- GenotypeStatsE10 %>% mutate(Range = paste(PeptideStart, PeptideEnd, sep = "-"))

p <- ggplot(data = GenotypeStatsE10, aes(x=Range, y=mean_val, fill = Genotype)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val), linewidth = 0.5, width = 0.2, position = position_dodge(width=0.9)) +
  scale_fill_manual(values=c("cyan4", "indianred", "brown")) + 
  theme_classic() +
  ylab("Peptide Abundance\n(standardised against the mean of constitutive peptides)") +
  xlab("MAPT Exon 10 Peptides (2N4R tau amino acid positions)")
p

ggsave_optional("Unscaled E10 only barplot.png")

```

Repeating plotting but with peptide-wise scaled data
```{r, fig.width=10,fig.height=6.5}
PeptideWiseMelt <- melt(PeptideWiseScaled, idvars = "Stripped.Sequence", measure.vars = as.character(c(1:20)))
names(PeptideWiseMelt)[names(PeptideWiseMelt) == "variable"] <- "ID"
head(PeptideWiseMelt)

#Add Genotype column
mapper$ID <- as.factor(mapper$ID)
PeptideWiseMelt <- left_join(PeptideWiseMelt, mapper[,colnames(mapper) %in% c("ID", "Genotype")], by = "ID")

#Genotype-wise stats
PeptideWiseGenotypeStats <- PeptideWiseMelt %>%
  group_by(Genotype, Stripped.Sequence) %>%
  summarise(
    count = n(),
    mean_val = mean(value, na.rm = TRUE),
    sd_val = sd(value, na.rm = TRUE)
  )

PeptideWiseGenotypeStats <- left_join(PeptideWiseGenotypeStats, PeptideWiseScaled[,1:5], by = "Stripped.Sequence")
head(PeptideWiseGenotypeStats)

#Amending IVS10+16 name for plotting
PeptideWiseGenotypeStats$Genotype <- gsub("IVS10_16", "IVS10+16", PeptideWiseGenotypeStats$Genotype)

```

Plotting
```{r}

#plotting - NB relies on determination of E10 box boundaries in a previous chunk
p <- ggplot(data=PeptideWiseGenotypeStats, aes(x=PeptideRank, y=mean_val, group = Genotype)) +
  geom_rect(ymax = Inf, ymin = -Inf, xmin = E10MinRank-0.5, xmax = E10MaxRank+0.5, fill = alpha("darkgoldenrod1", 0.01)) +
  geom_text(
    label="Exon 10", 
    x=mean(c(E10MinRank,E10MaxRank)),
    y=2.6,
    color = "#002000",
    aes(fontface=2),
    size = 4
  ) +
  geom_line(aes(color = Genotype), linewidth = 1.5, lineend = "round")+
  geom_point(aes(color = Genotype)) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val, color = Genotype), linewidth = 1, width = 0.4) +
  ggtitle("Peptides Ranked by Start Position") + 
  scale_color_manual(values=plot_colours) + 
  theme_classic() +
  ylab("Peptide Abundance\n(standardised individually)") +
  xlab("Peptide (ranked by start position)") +
  ylim(NA,2.5) +
  geom_hline(yintercept = 0, col = "black", linetype = 'dashed') +
  #facet_wrap to separate genotypes, remove to overlay as default
  facet_wrap(~Genotype, nrow=3)
print(p)

# ggsave_optional("Peptide-wise all peptides lineplot by rank facet_wrap.png")

p <- ggplot(data=PeptideWiseGenotypeStats, aes(x=PeptideRank, y=mean_val, group = Genotype)) +
  geom_rect(ymax = Inf, ymin = -Inf, xmin = E10MinRank-0.5, xmax = E10MaxRank+0.5, fill = alpha("darkgoldenrod1", 0.005)) +
  geom_text(
    label="Exon\n10", 
    x=mean(c(E10MinRank,E10MaxRank)),
    y=2.6,
    color = "#002000",
    aes(fontface=2),
    size = 4
  ) +
  geom_line(aes(color = Genotype), linewidth = 3)+
  geom_point(aes(color = Genotype)) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val, color = Genotype), linewidth = 0.5, width = 0.5) +
  # ggtitle("Peptides Ranked by Start Position") + 
  scale_color_manual(values=plot_colours) + 
  theme_classic() +
  ylab("Peptide Abundance\n(standardised within-peptide)") +
  xlab("MAPT Peptides (ranked by start position)") +
  ylim(NA,2.7) +
  geom_hline(yintercept = 0, col = "black", linetype = 'dashed') +
  theme(axis.title.x = element_text(size = 14, margin=margin(8,0,0,0)),
        axis.title.y = element_text(size = 14, margin=margin(0,8,0,0)),
        axis.text = element_text(size = 12, colour = "black"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        plot.title = element_text(size=16))
print(p)

ggsave_optional("Peptide-wise all peptides lineplot by rank.png")

q <- ggplot(data=PeptideWiseGenotypeStats, aes(x=PeptideStart, y=mean_val, group = Genotype)) +
  geom_rect(ymax = Inf, ymin = -Inf, xmin = E10MinPos-5, xmax = E10MaxPos+5, fill = alpha("darkgoldenrod1", 0.005)) +
  geom_text(
    label="Exon 10", 
    x=mean(c(E10MinPos, E10MaxPos)),
    y=2.6,
    color = "#002000",
    aes(fontface=2),
    size = 5
  ) +
  geom_line(aes(color = Genotype), linewidth = 1.5, lineend = "round")+
  geom_point(aes(color = Genotype)) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val, color = Genotype), linewidth = 1, width = 3) +
  ggtitle("Peptides by Actual Start Position") +
  scale_color_manual(values=plot_colours) +
  theme_classic() +
  ylab("Peptide Abundance\n(standardised individually)") +
  xlab("Peptides (actual start position)") +
  ylim(NA,2.5) +
  geom_hline(yintercept = 0, col = "black", linetype = 'dashed')
print(q)

# ggsave_optional("Peptide-wise all peptides lineplot by position.png")

```

Bar plot E10 only for PeptideScaled data
```{r}
PeptideWiseGenotypeStatsE10 <- PeptideWiseGenotypeStats[PeptideWiseGenotypeStats$InExon10,]
PeptideWiseGenotypeStatsE10 <- PeptideWiseGenotypeStatsE10 %>% mutate(Range = paste(PeptideStart, PeptideEnd, sep = "-"))

p <- ggplot(data = PeptideWiseGenotypeStatsE10, aes(x=Range, y=mean_val, fill = Genotype)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean_val - sd_val, ymax = mean_val + sd_val), linewidth = 0.5, width = 0.2, position = position_dodge(width=0.9)) +
  scale_fill_manual(values=c("cyan4", "indianred", "brown")) + 
  theme_classic() +
  ylab("Peptide Abundance\n(standardised individually)") +
  xlab("MAPT Exon 10 Peptides (2N4R tau amino acid positions)") +
  geom_hline(yintercept = 0, col = "black", linetype = 'dashed')
p

ggsave_optional("Peptide-wise E10 only barplot.png")
```


