---
title: "tPx plotting"
output: html_document
date: "2026-05-08"
---

```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
#library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
library(viridis)
library(ggrepel)
set.seed(42)
```


Parameters
```{r}
save_option <- TRUE


```

Load data and reformat
```{r}
three_genotype_file <- "./tPx Intermediate Outputs/three_genotype_imputed exclusions-TRUE NAThreshold-0.4 NonImputedKept-FALSE.csv"
tPx_data <- read.csv(three_genotype_file)

rownames(tPx_data) <- tPx_data$X
tPx_data <- tPx_data[,-1]

mapper <- read.csv("totalproteome_mapper.csv")
colnames(mapper)[which(names(mapper) == "Raw")] <- "DRI_ID"


save_id <- gsub(".*three_genotype_imputed ", "", three_genotype_file)
save_id <- gsub(".csv", "", save_id)
```

Filter out proteins with missing values
```{r}
complete_data <- as.data.frame(tPx_data)
complete_data[complete_data == 1] <- NA #not needed in final analysis as no missing values are kept so have never needed to be assigned values of 1
complete_data <- complete_data[complete.cases(complete_data),]
```

Normalise all data (NB probably not what Beth's pipeline does so not comparable)

NB this is standardising the data for each feature (i.e protein) so that they contribute equally to PCA weighting. The data being used here is the processed_normalised_filtered out of Beth's pipeline - it is already filtered, normalised (i.e. between samples to account for sample-sample variation in total coverage), and had imputation applied
```{r}
complete_data_scaled <- as.data.frame(scale(t(complete_data)))
#NB transforming as this is what PC will expect
```

Run PCA
```{r}
##Run PCA
PCA_output <- prcomp(x = complete_data_scaled)
fviz_eig(PCA_output, ncp = 25)
ind <- get_pca_ind(PCA_output)
pca_data <- as.data.frame(ind$coord)
head(pca_data)[,1:5]

explained_variances <- as.data.frame(summary(PCA_output)$importance)[2,]*100
```
Bring in genotype data
```{r}
pca_data$DRI_ID <- rownames(pca_data)
mapper$DRI_ID <- as.character(mapper$DRI_ID)
pca_data <- left_join(pca_data, mapper[,colnames(mapper) %in% c("DRI_ID", "Genotype")], by = "DRI_ID")
```

Plot PCA plots
```{r}
#Ensure correct IVS10+16 labelling
pca_data$Genotype <- gsub("IVS10_16", "IVS10+16", pca_data$Genotype)

plot_colours <- c("cyan4", "indianred", "darkred")

# Plot neat PCA plots with focus on Genotype comparison and clustering
for(base_pc in c(1,3,5,7,9,11)){
  second_pc <- base_pc + 1

  p <- ggplot(pca_data, aes(x=pca_data[,base_pc], y=pca_data[,second_pc], color=Genotype, label = DRI_ID)) +
    geom_point(size = 3) +
    stat_ellipse(data = pca_data, aes(x=pca_data[,base_pc], y=pca_data[,second_pc], color=Genotype, fill=Genotype), inherit.aes = FALSE, geom = "polygon", alpha=0.05, linewidth = 0.6) +
    xlab(paste0("PC", base_pc, ": ", explained_variances[1,base_pc], "% variance")) + 
    ylab(paste0("PC", second_pc, ": ", explained_variances[1,second_pc], "% variance")) + 
    # coord_fixed(ratio = 1) +
    ggtitle(paste0("PC", base_pc, " and PC", second_pc)) +
    # geom_text_repel(colour = "black", size = 2) +
    scale_color_manual(values = plot_colours) +
    scale_fill_manual(values = plot_colours) +
    theme(axis.title.x = element_text(size = 18, margin=margin(8,0,0,0)),
          axis.title.y = element_text(size = 18, margin=margin(0,8,0,0)),
          axis.text = element_text(size = 15, colour = "black"),
          legend.text = element_text(size = 12),
          legend.title = element_text(size = 12), 
          plot.title = element_text(size=20))
  print(p)
  if (save_option){
    ggsave(paste0("./tPx PCA plots/PC", base_pc, " vs ", second_pc, " ", save_id, ".png"))
  }
}
```


