---
title: "Data processing"
output: html_document
date: "2026-05-15"
---

This is all working off of the scripts provided by Beth in 'Beth Scripts/'

Load packages
```{r}
library(readxl)
library(readr)
library(reshape2)
library(ggplot2)
library(limma)
library(pheatmap)
library(mice)
library(preprocessCore)
library(corrplot)
library(odbc)
library(DBI)
library(RPostgres)
library(dplyr)
library(Rtsne)
library(rrcovNA)
library(tibble)

#source("~/Scripts/BETH_functions.r") #unknown if these are needed
```

Set optional settings
```{r}
set.seed(42)

plot_colours <- c("cyan4", "darkgoldenrod3")

remove_multi_ids <- TRUE #v2 - when TRUE, this now retains the MAPT data despite it's presence as a 'multi-id' protein

#max proportion of per-protein missingness allowed within a genotype for imputation to be attempted; proteins with higher missingness than this are not imputed
#0.35 allows 2 missing values per genotype when N=6-8. With exclusion, e.g. N=5 for one genotype, then 0.4 allows 2 missing values still, but now allows an N=8 genotype to have 3 missing values
imputation_missingness_threshold <- 0.4

#parameter on whether to include proteins excluded from imputation due to missingness in the output (don't need if only analysing proteins with 100% non-missing data, as all of these proteins will have missing data above the imputation_missingness_threshold proportion)
output_nonimputed_too <- FALSE

#DRI_ID of outlier samples to exclude, and option to exclude or not
exclude_outliers <- FALSE
outliers <- c("")

#save limma outputs?
save_limma <- TRUE

#save all_genotype_imputed intermediate data?
save_imputed <- TRUE

#save excluded proteins' data specifically?
save_excluded <- FALSE
```

========================================================================================
Notes on replacing variable names with more understandable replacements:
Aonly = report
m.Aonly, X20220915 = report_mat
norm.dat = report_mat_norm
Raw = DRI_ID
Condition = Conditions
types, sample.types, saved.con = all seem to be copies of Conditions
zeroProp = imputation_missingness_threshold


========================================================================================

Load tPx report matrix
```{r}
report <- read_delim("RN-MF_Set6_TP_ModelAD-RIKEN_20260730_report.pg_matrix.tsv", delim = "\t", escape_double = FALSE, trim_ws = TRUE)

#To remove proteins with multi-IDs separated by ";". v2 - THIS NOW RETAINS MAPT SPECIFICALLY
if(remove_multi_ids){
  report_MAPT <- report[grep("P10637", report$Protein.Group),]
  report <- report[grep(";",report$Protein.Group,invert = T),]
  report <- rbind(report_MAPT, report)
  report <- report[grep("P10637Mut",report$Protein.Group,invert = T),] # needed to exclude some 'SN' and 'PS' mutant taus also searched for
}


```

Reformatting report
```{r}
# Removing excessive filenaming
colnames(report) <- gsub(".*DIA","DRI",colnames(report))
colnames(report) <- gsub("_2025.*","",colnames(report))
colnames(report) <- gsub(".raw","",colnames(report))
# colnames(report) <- gsub(".2024.*","",colnames(report))
# colnames(report) <- gsub("_202.*","",colnames(report))

#Creating a data matrix from the report that excludes all non-data columns
report_mat <- data.matrix(report[,-c(1:6)])
rownames(report_mat) <- gsub(";.*","",report$Protein.Group) #NB even if multi-IDs still present, this names report_mat rows only by the first

head(report_mat)[,1:5]
```

Exclude any outliers?
```{r}
if (exclude_outliers){
  report_mat <- report_mat[,!colnames(report_mat) %in% outliers]
}
```

Load mapper
```{r}
mapper <- read.csv("totalproteome_mapper.csv")
# mapper <- mapper[,-c(1:3)]
# colnames(mapper)[which(names(mapper) == "Raw")] <- "DRI_ID"

print("SAMPLE INFORMATION AND IDS")
print(mapper)
```
Filter report_mat by genotype
```{r}
report_mat <- report_mat[,colnames(report_mat) %in% mapper[mapper$Genotype %in% c("H1.0", "MAPTKI"),1]]
```


Find the genotypes of all samples in the report matrix from the mapper, and create lists of the sample columns for each genotype
```{r}
#Pull Genotypes out of mapper in the same order as the samples are already in in the report matrix
Conditions <- mapper$Genotype[match(colnames(report_mat),mapper$DRI_ID)]
print("Number of samples in each conditions (here, Genotype)")
table(Conditions)


H1_cols <- grep("H1.0", Conditions)
MAPTKI_cols <- grep("MAPTKI", Conditions)


H1_DRI_IDs <- colnames(report_mat)[H1_cols]
MAPTKI_DRI_IDs <- colnames(report_mat)[MAPTKI_cols]

print("H1 samples' DRI_IDs")
print(H1_DRI_IDs)
print("MAPTKI samples' DRI_IDs")
print(MAPTKI_DRI_IDs)


```

Assessment of coverage per sample
```{r}
detected_per_sample <- as.data.frame(colSums(!is.na(report_mat)))
colnames(detected_per_sample) <- "Number.Detected.Proteins"
detected_per_sample$DRI_ID <- rownames(detected_per_sample)

p <- ggplot(detected_per_sample, aes(x=DRI_ID, y = Number.Detected.Proteins, group = Conditions, fill = Conditions)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = plot_colours) +
  ggtitle("Total Detected Proteins Per Sample") +
  ylab("Detected Proteins")
p


total_data_per_sample <- as.data.frame(colSums(report_mat, na.rm = TRUE))
colnames(total_data_per_sample) <- "Total.Data"
total_data_per_sample$DRI_ID <- rownames(total_data_per_sample)

p <- ggplot(total_data_per_sample, aes(x=DRI_ID, y = Total.Data, group = Conditions, fill = Conditions)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = plot_colours) +
  ggtitle("Total MS Signal Per Sample") +
  ylab("Total MS Signal")
p

########### GROUPING CONDITIONS TOGETHER

detected_per_sample <- left_join(detected_per_sample, mapper[,c("DRI_ID", "Genotype")], by = "DRI_ID")
detected_per_sample$Genotype[detected_per_sample$Genotype=="IVS10_16"] <- "IVS10+16"
detected_per_sample_summary <- detected_per_sample %>% group_by(Genotype) %>%
  summarise(
    count = n(),
    mean_val = mean(Number.Detected.Proteins, na.rm = TRUE),
    sd_val = sd(Number.Detected.Proteins, na.rm = TRUE)
  )

#name amendments
detected_per_sample_summary$Genotype <- gsub("H1.0", "MAPT-GR H1.0", detected_per_sample_summary$Genotype)
detected_per_sample_summary$Genotype <- gsub("MAPTKI", "MAPTKI WT", detected_per_sample_summary$Genotype)
detected_per_sample$Genotype <- gsub("H1.0", "MAPT-GR H1.0", detected_per_sample$Genotype)
detected_per_sample$Genotype <- gsub("MAPTKI", "MAPTKI WT", detected_per_sample$Genotype)


p <- ggplot() +
  geom_bar(aes(x=Genotype, y=mean_val, fill = Genotype), detected_per_sample_summary, stat = 'identity', width = 0.8) +
  geom_errorbar(aes(x=Genotype, ymin=mean_val-sd_val, ymax=mean_val+sd_val), detected_per_sample_summary, width = 0.2) +
  geom_jitter(aes(x=Genotype, y=Number.Detected.Proteins), detected_per_sample, size = 2, width = 0.3) +
  labs(y = "Protein Detections") +
  scale_fill_manual(values = plot_colours) +
  guides(fill = "none") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 18, margin=margin(0,10,0,0)),
        axis.text = element_text(size = 15, colour = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1))
p

ggsave("./tPx QC Plots/Total Detected Proteins by genotype.png", create.dir = TRUE, width = 3, height = 4.5)

```

Assessing NAs per detected protein
```{r}
rownas <- as.data.frame(rowSums(is.na(report_mat)))
rownas$Protein.Group <- rownames(report_mat)
colnames(rownas)[1] <- "Number.NAs"

rownas$Number.Not.Missing <- (ncol(report_mat) - rownas$Number.NAs)

ggplot(rownas, aes(Number.Not.Missing))+
  geom_bar(fill = "orange")+
  ylab("Number of proteins")+
  xlab("Number of samples detected in")+
  ggtitle("Histogram of per-protein coverage")+
  theme_bw()

print("Number of proteins with x missing values")
table(rownas$Number.NAs)

print("Number of proteins with x values present")
table(rownas$Number.Not.Missing)
```

Data normalisation
```{r}
#normalizeMedianValues adjusts each column to have the same median value. NB this function log2 transforms the data, subtracts the difference between the sample median and the global median for each sample, and then un-transforms the data back to the original scale
report_mat_norm <- as.data.frame(normalizeMedianValues(report_mat))

report_mat_norm$Protein.Group <- rownames(report_mat_norm)

report_mat_norm_melted <- melt(report_mat_norm)
report_mat_melted <- melt(report_mat)

report_mat_norm_melted$Norm <- "Normalised"
report_mat_melted$Norm <- "Not Normalised"

colnames(report_mat_melted) <- colnames(report_mat_norm_melted)

norm_vs_non <- rbind(report_mat_norm_melted,report_mat_melted)
norm_vs_non$value <- log(norm_vs_non$value,2)

ggplot(norm_vs_non,aes(value,variable, fill = Norm))+
  facet_grid(cols=vars(Norm))+
  scale_fill_manual(values = c('#4363d8', '#f58231'))+
  xlab("")+
  ylab("")+
  geom_boxplot()+
  theme_bw()
```

Imputation - splitcheck. This means that each genotype is assessed for missingness separately and then imputed separately, to ensure that A) where data for a given genotype is imputed, it is not influenced by other genotypes, and B) when this happens, the given genotype has sufficient data for decent imputation
```{r}
#v2 - log transform report_mat_norm BEFORE imputation
report_mat_norm <- log2(report_mat_norm[,grep("DRI_", colnames(report_mat_norm))])

#Set threshold for imputation (only if not already done globally at the top of the script)
#imputation_missingness_threshold <- 0.35
n_proteins_before <- nrow(report_mat_norm)

########## H1 ##########

#Isolate H1 data columns
H1_mat <- report_mat_norm[,H1_cols]

#Determine which rows to keep, based on the (per-row number of NAs)/(number of columns) being less than the threshold
keep_rows <- rowSums(is.na(H1_mat))/ncol(H1_mat) <= imputation_missingness_threshold

#Filter H1_mat for these rows
H1_mat <- H1_mat[keep_rows,]

#Summarise
H1_num_proteins_kept <- nrow(H1_mat)
print("FOR H1 GENOTYPE")
print(paste0(H1_num_proteins_kept," kept, out of ", n_proteins_before, " (i.e. for which less than ",imputation_missingness_threshold*100,"% of samples have missing data)"))

#Extract what proteins were filtered out of analysis
cut_rows <- 1:nrow(report_mat_norm) #make a sequence that corresponds to a list of row numbers in the original report_mat
cut_rows <- cut_rows[!cut_rows %in% match(rownames(H1_mat),rownames(report_mat_norm))] #filter out any rows numbers for which the corresponding protein in report_mat is still present in the filtered H1_mat
cut_mat <- report_mat_norm[cut_rows,H1_cols] #create a new cut_mat, which has only the genotype's data columns and only the rows of filtered-out proteins (too high missingness for imputation)

if(save_excluded){
  write.csv(cut_mat, file = "./tPx Intermediate Outputs/cut_mat of proteins excluded from H1 imputation.csv")
}
#Imputation!
H1_imputed <- impSeqRob(H1_mat)$x

#Combine with non-imputed data, if desired
if (output_nonimputed_too){
  H1_mat <- rbind(H1_mat, cut_mat)
}

########## MAPTKI ##########

#Isolate MAPTKI data columns
MAPTKI_mat <- report_mat_norm[,MAPTKI_cols]

#Determine which rows to keep, based on the (per-row number of NAs)/(number of columns) being less than the threshold
keep_rows <- rowSums(is.na(MAPTKI_mat))/ncol(MAPTKI_mat) <= imputation_missingness_threshold

#Filter MAPTKI_mat for these rows
MAPTKI_mat <- MAPTKI_mat[keep_rows,]

#Summarise
MAPTKI_num_proteins_kept <- nrow(MAPTKI_mat)
print("FOR MAPTKI GENOTYPE")
print(paste0(MAPTKI_num_proteins_kept," kept, out of ", n_proteins_before, " (i.e. for which less than ",imputation_missingness_threshold*100,"% of samples have missing data)"))

#Extract what proteins were filtered out of analysis
cut_rows <- 1:nrow(report_mat_norm) #make a sequence that corresponds to a list of row numbers in the original report_mat_norm
cut_rows <- cut_rows[!cut_rows %in% match(rownames(MAPTKI_mat),rownames(report_mat_norm))] #filter out any rows numbers for which the corresponding protein in report_mat_norm is still present in the filtered MAPTKI_mat
cut_mat <- report_mat_norm[cut_rows,MAPTKI_cols] #create a new cut_mat, which has only the genotype's data columns and only the rows of filtered-out proteins (too high missingness for imputation)

if(save_excluded){
  write.csv(cut_mat, file = "./tPx Intermediate Outputs/cut_mat of proteins excluded from MAPTKI imputation.csv")
}

#Imputation!
MAPTKI_imputed <- impSeqRob(MAPTKI_mat)$x

#Combine with non-imputed data, if desired
if (output_nonimputed_too){
  MAPTKI_mat <- rbind(MAPTKI_mat, cut_mat)
}
```

Full join all genotypes back together (NB need to first copy rownames to a column to be accessible by full_join)
NOTE that the final df's columns will no longer be in the same order as the original report_mat
```{r}
#full joins
all_genotype_imputed <- full_join(rownames_to_column(as.data.frame(H1_imputed)), rownames_to_column(as.data.frame(MAPTKI_imputed)), by = "rowname")
#move rownames back to the df rownames() and remove from the df data itself, for neatness
rownames(all_genotype_imputed) <- all_genotype_imputed$rowname
all_genotype_imputed <- all_genotype_imputed %>% select(!rowname)

#save to tPx Intermediate Outputs folder with relevant parameters recorded in the filename
if(save_imputed){
  write.csv(all_genotype_imputed, paste0("./tPx Intermediate Outputs/all_genotype_imputed exclusions-", exclude_outliers, " NAThreshold-", imputation_missingness_threshold, " NonImputedKept-", output_nonimputed_too, ".csv"))
}
```

Assess negative values (these will fail to go through log2 transformation later)
```{r}
negative_imputed_rows <- all_genotype_imputed[rowSums(all_genotype_imputed < -1, na.rm = TRUE)>0,]

print("Number of rows with any negative values produced by imputation")
print(nrow(negative_imputed_rows))
```


======================== DIFFERENTIAL ANALYSIS ===============================
NB only handling data with 100% present data

```{r}
#Log transform the data and convert all introduced NaN values to NA ---- v2 no longer needed as data is logged before imputation
# all_genotype_imputed_log <- suppressWarnings(log2(all_genotype_imputed))
# all_genotype_imputed_log[all_genotype_imputed_log=="NaN"] <- NA
all_genotype_imputed_log <- all_genotype_imputed

########## H1 vs MAPTKI ###########
#specify pairwise conditions - control first!
pairwise_cons <- c("H1.0", "MAPTKI")

#selected pairwise data and filter out any rows containing any missing data
pairwise_H1_MAPTKI <- all_genotype_imputed_log[,c(H1_DRI_IDs, MAPTKI_DRI_IDs)]
pairwise_H1_MAPTKI <- pairwise_H1_MAPTKI[complete.cases(pairwise_H1_MAPTKI),]

#prepare vector defining Case vs Control, corresponding th the new column order in the filtered pairwise data
pairwise_cons_long <- c(rep(pairwise_cons[1], length(H1_DRI_IDs)), rep(pairwise_cons[2], length(MAPTKI_DRI_IDs)))
pairwise_cons_long[pairwise_cons_long==pairwise_cons[1]] <- "Control"
pairwise_cons_long[pairwise_cons_long==pairwise_cons[2]] <- "Case"


#limma
design <- model.matrix(~ -1+factor(pairwise_cons_long))
colnames(design) <- c("Control","Case")
contrast.matrix <- makeContrasts(Control-Case,levels=design)
fit <- lmFit(pairwise_H1_MAPTKI,design)
fit2 <- contrasts.fit(fit, contrast.matrix)
efit2 <- eBayes(fit2)

#pull out and sort data
pairwise_H1_MAPTKI_results <- topTreat(efit2, coef=1, adjust="fdr",sort.by="logFC",number=50000)
pairwise_H1_MAPTKI_results <- pairwise_H1_MAPTKI_results[order(pairwise_H1_MAPTKI_results$P.Value, decreasing = FALSE),]

#number of significant results a=0.05
print(paste0("Comparison of ", pairwise_cons[1], " and ", pairwise_cons[2]))
print("Number of significant differential proteins/pSites based on P.Value < 0.05")
nrow(pairwise_H1_MAPTKI_results %>% filter(P.Value<0.05))

#save
if(save_limma){
  write.csv(pairwise_H1_MAPTKI_results, paste0("./tPx Final Outputs/",
                                            pairwise_cons[1],
                                            " vs ",
                                            pairwise_cons[2],
                                            " exclusions-", exclude_outliers, 
                                            " NAThreshold-", imputation_missingness_threshold,
                                            ".csv"))
}
```


To do:
save limma outputs
try regressing out PC1 (whatever that means)
- including in limma model? Might need to understand more about how limma works for this
exclude sample 2 due to being an outlier?
- Actually for the tPx data, the batch effect appears in PC2!


To test including PC2 in the limma model, first run PCA
```{r}
#PCA
# library(FactoMineR)
# library(factoextra)
# library(ggrepel)
# 
# pca_input <- all_genotype_imputed[complete.cases(all_genotype_imputed),]
# complete_data_scaled <- as.data.frame(scale(t(pca_input)))
# pca_output <- prcomp(x = complete_data_scaled)
# ind <- get_pca_ind(pca_output)
# pca_data <- as.data.frame(ind$coord)
# head(pca_data)[,1:5]
# explained_variances <- as.data.frame(summary(pca_output)$importance)[2,]*100
# 
# pca_data$DRI_ID <- rownames(pca_data)
# mapper$DRI_ID <- as.character(mapper$DRI_ID)
# pca_data <- left_join(pca_data, mapper[,colnames(mapper) %in% c("DRI_ID", "Genotype")], by = "DRI_ID")
```

Plot the PCA plots
```{r}
# Plot neat PCA plots with focus on Genotype comparison and clustering
# for(base_pc in c(1,3,5,7,9,11)){
#   second_pc <- base_pc + 1
# 
#   p <- ggplot(pca_data, aes(x=pca_data[,base_pc], y=pca_data[,second_pc], color=Genotype, label=DRI_ID)) +
#     geom_point() +
#     geom_text_repel(size = 2, color = "black", force_pull = 2,) +
#     stat_ellipse(data = pca_data, aes(x=pca_data[,base_pc], y=pca_data[,second_pc], color=Genotype, fill=Genotype), inherit.aes = FALSE, geom = "polygon", alpha=0.15) +
#     xlab(paste0("PC", base_pc, ": ", explained_variances[1,base_pc], "% variance")) + 
#     ylab(paste0("PC", second_pc, ": ", explained_variances[1,second_pc], "% variance")) + 
#     coord_fixed() +
#     ggtitle(paste0("PC", base_pc, " and PC", second_pc))
#   print(p)
# }
# 
# ##for use in limma
# chosen_pc <- "Dim.1"
# pc_to_use_as_covariate <- pca_data[,c(chosen_pc, "DRI_ID", "Genotype")]
```

v2 - PCA outlier detection
```{r}
# outlier_analysis <- mt::pca.outlier(t(pca_input), xlim = c(-160, 160), ylim = c(-100, 100), conf.level = 0.94, groups = mapper$Genotype[match(colnames(pca_input), mapper$DRI_ID)])
# print(outlier_analysis$plot)
# 
# 
# print(paste0(length(outlier_analysis$outlier), " samples reach a 95% confidence cutoff for outliers (NB the default is 0.975, so even more stringent)"))
# 
# if(length(outlier_analysis$outlier)>0){
#   print(paste("Outliers:", colnames(pca_input)[outlier_analysis$outlier], sep = " "))
# }
```


v2 - no PC inclusion in limma output
```{r}
# ########## H1 vs IVS10+16 ###########
# #specify pairwise conditions - control first!
# pairwise_cons <- c("H1.0", "IVS10_16")
# 
# #selected pairwise data and filter out any rows containing any missing data
# pairwise_H1_IVS_PCout <- all_genotype_imputed_log[,c(H1_DRI_IDs, IVS_DRI_IDs)]
# pairwise_H1_IVS_PCout <- pairwise_H1_IVS_PCout[complete.cases(pairwise_H1_IVS_PCout),]
# 
# #prepare vector defining Case vs Control, corresponding th the new column order in the filtered pairwise data
# pairwise_cons_long <- c(rep(pairwise_cons[1], length(H1_DRI_IDs)), rep(pairwise_cons[2], length(IVS_DRI_IDs)))
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[1]] <- "Control"
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[2]] <- "Case"
# 
# 
# #limma
# design <- model.matrix(~ -1 + factor(pairwise_cons_long) + pc_to_use_as_covariate[pc_to_use_as_covariate$Genotype %in% pairwise_cons, chosen_pc])
# colnames(design) <- c("Control","Case", chosen_pc)
# contrast.matrix <- makeContrasts(Control-Case,levels=design)
# fit <- lmFit(pairwise_H1_IVS_PCout,design)
# fit2 <- contrasts.fit(fit, contrast.matrix)
# efit2 <- eBayes(fit2)
# 
# #pull out and sort data
# pairwise_H1_IVS_PCout_results <- topTreat(efit2, coef=1, adjust="fdr",sort.by="logFC",number=50000)
# pairwise_H1_IVS_PCout_results <- pairwise_H1_IVS_PCout_results[order(pairwise_H1_IVS_PCout_results$P.Value, decreasing = FALSE),]
# 
# #number of significant results a=0.05
# print(paste0("Comparison of ", pairwise_cons[1], " and ", pairwise_cons[2]))
# print("Number of significant differential proteins/pSites based on P.Value < 0.05")
# nrow(pairwise_H1_IVS_PCout_results %>% filter(P.Value<0.05))
# 
# if(save_limma){
#   write.csv(pairwise_H1_IVS_PCout_results, paste0("./tPx Final Outputs/",
#                                             pairwise_cons[1],
#                                             " vs ",
#                                             pairwise_cons[2],
#                                             " exclusions-", exclude_outliers, 
#                                             " NAThreshold-", imputation_missingness_threshold,
#                                             " ", chosen_pc,
#                                             " out",
#                                             ".csv"))
# }
# 
# ########## H1 vs N279K ###########
# #specify pairwise conditions - control first!
# pairwise_cons <- c("H1.0", "N279K")
# 
# #selected pairwise data and filter out any rows containing any missing data
# pairwise_H1_N279K_PCout <- all_genotype_imputed_log[,c(H1_DRI_IDs, N279K_DRI_IDs)]
# pairwise_H1_N279K_PCout <- pairwise_H1_N279K_PCout[complete.cases(pairwise_H1_N279K_PCout),]
# 
# #prepare vector defining Case vs Control, corresponding th the new column order in the filtered pairwise data
# pairwise_cons_long <- c(rep(pairwise_cons[1], length(H1_DRI_IDs)), rep(pairwise_cons[2], length(N279K_DRI_IDs)))
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[1]] <- "Control"
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[2]] <- "Case"
# 
# 
# #limma
# design <- model.matrix(~ -1 + factor(pairwise_cons_long) + pc_to_use_as_covariate[pc_to_use_as_covariate$Genotype %in% pairwise_cons, chosen_pc])
# colnames(design) <- c("Control","Case", chosen_pc)
# contrast.matrix <- makeContrasts(Control-Case,levels=design)
# fit <- lmFit(pairwise_H1_N279K_PCout,design)
# fit2 <- contrasts.fit(fit, contrast.matrix)
# efit2 <- eBayes(fit2)
# 
# #pull out and sort data
# pairwise_H1_N279K_PCout_results <- topTreat(efit2, coef=1, adjust="fdr",sort.by="logFC",number=50000)
# pairwise_H1_N279K_PCout_results <- pairwise_H1_N279K_PCout_results[order(pairwise_H1_N279K_PCout_results$P.Value, decreasing = FALSE),]
# 
# #number of significant results a=0.05
# print(paste0("Comparison of ", pairwise_cons[1], " and ", pairwise_cons[2]))
# print("Number of significant differential proteins/pSites based on P.Value < 0.05")
# nrow(pairwise_H1_N279K_PCout_results %>% filter(P.Value<0.05))
# 
# if(save_limma){
#   write.csv(pairwise_H1_N279K_PCout_results, paste0("./tPx Final Outputs/",
#                                             pairwise_cons[1],
#                                             " vs ",
#                                             pairwise_cons[2],
#                                             " exclusions-", exclude_outliers, 
#                                             " NAThreshold-", imputation_missingness_threshold,
#                                             " ", chosen_pc,
#                                             " out",
#                                             ".csv"))
# }
# 
# ########## IVS10+16 vs N279K ###########
# #specify pairwise conditions - control first!
# pairwise_cons <- c("IVS10_16", "N279K")
# 
# #selected pairwise data and filter out any rows containing any missing data
# pairwise_IVS_N279K_PCout <- all_genotype_imputed_log[,c(IVS_DRI_IDs, N279K_DRI_IDs)]
# pairwise_IVS_N279K_PCout <- pairwise_IVS_N279K_PCout[complete.cases(pairwise_IVS_N279K_PCout),]
# 
# #prepare vector defining Case vs Control, corresponding to the new column order in the filtered pairwise data
# pairwise_cons_long <- c(rep(pairwise_cons[1], length(IVS_DRI_IDs)), rep(pairwise_cons[2], length(N279K_DRI_IDs)))
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[1]] <- "Control"
# pairwise_cons_long[pairwise_cons_long==pairwise_cons[2]] <- "Case"
# 
# 
# #limma
# design <- model.matrix(~ -1 + factor(pairwise_cons_long) + pc_to_use_as_covariate[pc_to_use_as_covariate$Genotype %in% pairwise_cons, chosen_pc])
# colnames(design) <- c("Control","Case", chosen_pc)
# contrast.matrix <- makeContrasts(Control-Case,levels=design)
# fit <- lmFit(pairwise_IVS_N279K_PCout,design)
# fit2 <- contrasts.fit(fit, contrast.matrix)
# efit2 <- eBayes(fit2)
# 
# #pull out and sort data
# pairwise_IVS_N279K_PCout_results <- topTreat(efit2, coef=1, adjust="fdr",sort.by="logFC",number=50000)
# pairwise_IVS_N279K_PCout_results <- pairwise_IVS_N279K_PCout_results[order(pairwise_IVS_N279K_PCout_results$P.Value, decreasing = FALSE),]
# 
# #number of significant results a=0.05
# print(paste0("Comparison of ", pairwise_cons[1], " and ", pairwise_cons[2]))
# print("Number of significant differential proteins/pSites based on P.Value < 0.05")
# nrow(pairwise_IVS_N279K_PCout_results %>% filter(P.Value<0.05))
# 
# if(save_limma){
#   write.csv(pairwise_IVS_N279K_PCout_results, paste0("./tPx Final Outputs/",
#                                             pairwise_cons[1],
#                                             " vs ",
#                                             pairwise_cons[2],
#                                             " exclusions-", exclude_outliers, 
#                                             " NAThreshold-", imputation_missingness_threshold,
#                                             " ", chosen_pc,
#                                             " out",
#                                             ".csv"))
#}
```



